Dissertation Proposal

### 2023-11-10 proposal.tex
group sections

%% abbrev. for quizzes
\newcommand{\QuizFun}{Quiz 1}
\newcommand{\QuizState}{Quiz 2}
\newcommand{\QuizHof}{Quiz 3}
\newcommand{\quizFun}{Quiz 1}
\newcommand{\quizState}{Quiz 2}
\newcommand{\quizHof}{Quiz 3}

@@

To overcome such bias, I analyze a University~1 dataset produced by a
two-year process to find (a) student-written programs that tend to
trip up students and (b) incorrect responses to those programs
(\Cref{s:gen-prob-quizius}), using a tool for the purpose
(\Cref{s:quizius}). I have been cleaning up those programs
(\Cref{s:smol-quizzes,s:tutor-refine-program-set}) and confirming that
these programs are still tricky in the same way
(\Cref{s:misconceptions-quiz,s:misconceptions-tutor}).
\Cref{s:misconceptions-tutor} presents the final list of
misconceptions found with this process.

@@

\section{Generating and Collating Problems}%
\label{s:gen-prob-quizius}

% Additionally, in this paper, I intentionally use the word
% mis\emph{conceptions} rather than mis\emph{take}. A mistake can happen
% for any reason (e.g., selecting the wrong answer from a menu). A
% misconception, however, implies a conceptual problem: the student has
% formed an incorrect concept in their head. For instance, they may
% think that mutable structures are copied on function calls, or that
% scope is resolved dynamically. This requires probing what they are
% thinking.

% All these, however, add up to a somewhat challenging demand. I want
% to produce a list of questions (each one an MCQ) such that
% \begin{enumerate}

%   \item I can get past the expert blind spot,

%   \item I have a sense of what misconceptions students have, and

%   \item I can generally associate wrong answers with specific
%         misconceptions, approaching a concept inventory.

This section describes preliminary results on finding misconceptions
from student-written programs (\Cref{s:quizius}) and what I have done
to distill misconception-detecting programs from student inputs
(\Cref{s:smol-quizzes,s:tutor-refine-program-set}).

@@

My main solution to the expert blind spot is to use the Quizius system
\citep{saarinenHarnessingWisdomClasses2019}. In contrast to the very
heavyweight process (involving a lot of expert time) that is generally
used to create a concept inventory, Quizius uses a lightweight,
interactive approach to obtain fairly comparable data, which an expert
can then shape into a quality instrument.

@@

\subsection{Collating Problems}
\label{s:smol-quizzes}

@@

\end{itemize}
and so on.
I therefore manually curated the Quizius output
to address these issues.

% \begin{figure}[t]
%   \centering
%   \includegraphics[width=\linewidth]{images/smol-quizzes-example.png}
%   \caption{Screenshot of a SMoL Quiz question.}
%   \label{f:smol-quizzes-example}
% \end{figure}

\subsection{The SMoL Quizzes}%
\label{s:misconceptions-quiz}

Having curated the output, I had to confirm that these programs were
still effective! That is, they needed to actually find student errors.

I therefore turned the programs into a set of quizzes (in the US
sense: namely, a brief test of knowledge) that I call the SMoL
Quizzes.
% There were three quizzes, ordered by linguistic
% complexity. The first consisted of only basic operators and
% first-order functions. The second added variable and structure
% mutation. The third added \lstinline|lambda| and higher-order functions.
Each quiz is a sequence of MCQs. The options are from the Quizius
data. In addition, there is an ``Other'' option. If chosen, the quiz
gave the user a text box with the caption ``Please specify''. The goal
here was to record any other answers, which in turn might lead to
fresh misconceptions.

The quizzes were administered in the third year in University 1.

% Question orders were partially randomized. I wanted students to get
% some easy, warm-up questions initially, so those were kept at the
% beginning. Similarly, I wanted programs that are syntactically
% similar to stay close to each other in the quiz. This is so that, when
% students got a second such program, they would not have to look far to
% find the first one and confirm that they are indeed (slightly)
% different, rather than wonder if they were seeing the same program
% again.

% Students only received feedback after having completed a whole quiz.
% At the end of each quiz, they received both summative feedback
% \emph{and} a refutation text that explained every program that
% appeared in the quiz. Students were also encouraged to run the
% programs, but I have little reason to believe that they did (and
% certainly they asked few questions on the class forum about them).

% Due to space limitations, I present the entire instrument in
% \supplRef{SMoL Quizzes/instrument}. Here I focus on a few programs
% where student choices correspond to misconceptions identified earlier,
% thereby also showing that the curated programs are still effective. I
% use a * to indicate the correct answer (which also matches the
% implementation's output).

% Syntactically, \lstinline|#t| and \lstinline|#f| are true and false,
% while \lstinline|#(...)| is a vector.

% This quiz starts with two warm-up and not very interesting questions
% and then other questions. The order of the other questions are fully
% randomized.

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{l X}
    \textbf{Questions}                       &
    \textbf{Table of Answers}                    \\\hline
    \lstinputlisting{programs/def_twice.txt} & {
        \begin{tabularx}{\linewidth}{r X}
          $^\dag$59\% & \lstinline|2 0|             \\
          *32\%       & \lstinline|Error|           \\
          4\%         & \lstinline|Other: 2 2|      \\
          3\%         & ``Nothing is printed'' \\
          2\%         & \lstinline|0 0|             \\
        \end{tabularx}
    }                                            \\\hline
    \lstinputlisting{programs/dyn_scope.txt} & {
        \begin{tabularx}{\linewidth}{r X}
          *65\%       & \lstinline|Error| \\
          $^\dag$29\% & \lstinline|42 #t| \\
          6\%         & Other        \\
        \end{tabularx}
    }                                            \\\hline
    % \lstinputlisting{programs/reassign.txt}  & {
    %     \begin{tabularx}{\linewidth}{r X}
    %       *53\%       & \lstinline|#(6 5) 5|                 \\
    %       *20\%       & \lstinline|Error|                    \\
    %       $^\dag$11\% & \lstinline|#(6 6) 5|                 \\
    %       $^\dag$11\% & \lstinline|#(6 6) 6|                 \\
    %       5\%         & ``Nothing is printed'' or Other \\
    %     \end{tabularx}
    % }                                            \\\hline
  \end{tabularx}
  \caption{Some questions and student answers from \quizFun. Each
    table row is a program and the (relative) frequencies of
    student answers. Answers that can be considered correct are marked
    with \texttt{*}.}
  \label{t:quizzes-data-fun}
\end{table}
\Cref{t:quizzes-data-fun} lists programs in \quizFun{} that I
consider the most interesting. These data confirm the presence of
scope-related misconceptions.
% :
% \begin{enumerate}
%   \item At least 59\% of students incorrectly believe that a
%         variable can be defined twice in one block.
%   \item 29\% of students believe that a variable defined in a function
%         will be available in the top-level (or global) environment
%         after the function is called. That is, these students may have a
%         dynamic scope misconception.
%   \item 22\% (11\% + 11\%) of students believe variables themselves
%         can be passed as arguments and redefined inside the function.
%         They disagree on whether the redefinition persists after the
%         function call.
% \end{enumerate}

% This quiz randomizes the order of all questions.

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{l X}
    \textbf{Question}                           &
    \textbf{Table of Answers}                       \\\hline
    \lstinputlisting{programs/circularity.txt}  & {
        \begin{tabularx}{\linewidth}{r X}
          $^\dag$50\% & \lstinline|#(#(2 #(2 3)) #(2 3))| \\
          *29\%       & \lstinline|x=#(x x)| or
          something similar. Both \lstinline|(left x)| and \lstinline|(right x)|
          are \lstinline|x| itself.                       \\
          $^\dag$16\% & \lstinline|Error|                 \\
          5\%         & Other                        \\
        \end{tabularx}
    }                                               \\\hline
    % \lstinputlisting{programs/mvec_in_mvec.txt} & {
    %     \begin{tabularx}{\linewidth}{r X}
    %       *62\%       & \lstinline|#(#(100 2 3 4) #(100 2 3 4))| \\
    %       $^\dag$26\% & \lstinline|#(#(1 2 3 4) #(100 2 3 4))|   \\
    %       6\%         & \texttt{Error}                      \\
    %       4\%         & \lstinline|#(#(1 2 3 4) #(1 2 3 4))|     \\
    %       2\%         & Other                               \\
    %     \end{tabularx}
    % }                                               \\\hline
    % \lstinputlisting{programs/vset_in_let.txt}  & {
    %     \begin{tabularx}{\linewidth}{r X}
    %       *59\%       & \lstinline|#(10)|  \\
    %       $^\dag$36\% & \lstinline|#(123)| \\
    %       5\%         & Other         \\
    %     \end{tabularx}
    % }                                               \\\hline
    \lstinputlisting{programs/var_as_arg.txt}   & {
        \begin{tabularx}{\linewidth}{r X}
          *65\%       & \texttt{12} \\
          $^\dag$31\% & \texttt{0}  \\
          4\%         & Other       \\
        \end{tabularx}
    }                                               \\\hline
    % \lstinputlisting{programs/setset.txt}       & {
    %     \begin{tabularx}{\linewidth}{r X}
    %       *59\%       & Other: \texttt{5 7}   \\
    %       $^\dag$27\% & \texttt{6 7}          \\
    %       11\%        & \texttt{5 5}          \\
    %       2\%         & Other: \texttt{Error} \\
    %       1\%         & Other: \texttt{5 6}   \\
    %     \end{tabularx}
    % }                                               \\\hline
  \end{tabularx}
  \caption{Some questions and student answers from \quizState. Each
    table row is a program and the (relative) frequencies of
    student answers. Answers that can be considered correct are marked
    with \texttt{*}.}
  \label{t:quizzes-data-state}
\end{table}
\Cref{t:quizzes-data-state} lists interesting programs after the
addition of state. These data suggest the students have aliasing-related misconceptions.
% :
% \begin{itemize}
%   \item Up to 50\% of students think vectors are copied rather than
%         aliased.
%   \item 16\% of students think trying to construct and print a
%         self-referring vector would error. (This program is ambiguous:
%         \emph{constructing} works in most SMoL languages, but \emph{printing}
%         may well cause a problem. These identified ambiguities are
%         addressed in \Cref{s:tutor-refine-program-set}.)
%   \item 31\% of students think a variable is aliased by a parameter if
%         the two variables have the same name. Perhaps interestingly,
%         fewer students (27\%) think the variable aliasing would
%         happen if the two variables had different names.
% \end{itemize}

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{l X}
    \textbf{Question}                                 &
    \textbf{Table of Answers}                             \\\hline
    \lstinputlisting{programs/fun_returns_lambda.txt} & {
        \begin{tabularx}{\linewidth}{r X}
          *82\%       & \lstinline|3|     \\
          $^\dag$17\% & \lstinline|Error| \\
          1\%         & Other        \\
        \end{tabularx}
    }                                                     \\\hline
    \lstinputlisting{programs/fun_and_state_1.txt}    & {
        \begin{tabularx}{\linewidth}{r X}
          *74\%       & \lstinline|4|            \\
          $^\dag$21\% & \lstinline|3|            \\
          6\%         & Other: \lstinline|Error| \\
        \end{tabularx}
    }                                                     \\\hline
    % \lstinputlisting{programs/fun_and_state_2.txt}    & {
    %     \begin{tabularx}{\linewidth}{r X}
    %       *88\%       & \lstinline|4|            \\
    %       $^\dag$10\% & \lstinline|3|            \\
    %       3\%         & Other: \lstinline|Error| \\
    %     \end{tabularx}
    % }                                                     \\\hline
    \lstinputlisting{programs/counter.txt}            & {
        \begin{tabularx}{\linewidth}{r X}
          *62\%       & \lstinline|1 1 2 3 2| \\
          $^\dag$34\% & \lstinline|1 1 1 1 1| \\
          4\%         & Other            \\
          1\%         & \lstinline|1 1 2 3 4| \\
        \end{tabularx}
    }                                                     \\\hline
  \end{tabularx}
  \caption{Some questions and student answers from \quizHof. Each
    table row is a program and the (relative) frequencies of
    student answers. Answers that can be considered correct are marked
    with \texttt{*}.}
  \label{t:quizzes-data-hof}
\end{table}
\Cref{t:quizzes-data-hof} lists interesting programs after the
addition of closures and higher-order functions. It extends what we
saw with loops in \Cref{s:intro}: that students have misconceptions
about their interaction with mutable variables.

\subsection{Further Collating Problems, for the Tutor}%
\label{s:tutor-refine-program-set}

The SMoL Tutor's interpreting questions are the final instrument of
this paper. I bootstrap it from the SMoL Quizzes, but made the
following alterations:
\begin{enumerate}

  % fill gaps in goal sentences
  \item I felt the existing programs were insufficient to cover the
        ideas I wanted students to work through. For instance, SMoL
        Quizzes did not alias vectors using \lstinline|defvar|, so we
        added this program:

@@

\end{enumerate}
Most of these steps either modify or elide programs. Two cases,
filling gaps in light of the goal sentences and resolving ambiguities,
introduce programs. Starting with the 37 programs in the SMoL Quizzes,
I added 52 more programs to arrive at a total of 89.

In addition to generating this set of programs, I also modified the
answer options. In addition to retaining the correct and incorrect
answers from the SMoL Quizzes (and constructing my best guess of
analogous incorrect answers for the new problems), I added more wrong
answers to for the reasons given in \Cref{s:interpreting-tasks}. In
general, I ended up increasing the number of choices substantially:
only 8 out of the 89 questions have three or four choices; 70
questions have 5--8 choices; and 11 questions have 9--14 choices.

\section{Preliminary List of Misconceptions}%
\label{s:misconceptions-tutor}

@@

I do not present the results of and \textbf{local}
(\Cref{t:tutor-goal-sentences}), because they focus on constructs not
usually found in non-Lispy languages. They were included in the Tutor
to help students write programs for the course, but are not
interesting from a SMoL perspective.

% \subsection{Misinterpreters}
% \label{s:misinterp-intro}

% In principle, an expert can identify what misconceptions a particular
% incorrect answer might correspond to. In practice, I found this
% rather difficult for three reasons. First, with 89 programs, each with
% several answers, it's easy to make mistakes. Second, my own expert blind
% spot may prevent me from seeing an interpretation that would
% lead to an additional association. Finally, and specific to my case,
% since I had either curated or written all the programs and answers,
% I was likely to miss some associations I had not intended.

% To address this problem, I formalized misconceptions as
% interpreters. That is, for each misconception I created a
% corresponding \emph{misinterpreter}. This is an interpreter for the
% SMoL syntax that intentionally has a semantic error corresponding to
% that misconception. Put differently, a misinterpreter is a like a
% ``definitional interpreter for a misconception''.

% By running programs through the misinterpreters, I can more uniformly
% and rigorously identify \emph{all} the misconceptions associated with
% a wrong answer. Furthermore, if I identify a new misconception (or
% alter a misinterpreter), it is easy to automatically re-classify all
% the answers. By using misinterpreters, I indeed found new
% interpretations for existing program-answer pairs.

% \subsection{A Catalog of Misconceptions}

I iteratively created our final catalog of misconceptions (from the
perspective of the data in this paper). I started with misinterpreters
representing the misconceptions described in
\Cref{s:misconceptions-quiz}. These are misconceptions for which we
have reasonable validation (due to the prose in Quizius), so I call
them \emph{grounded} misconceptions. I then looked at wrong answers
not covered by these but chosen by students, and did my best to
distill these into misconceptions. These are \emph{surmised}
misconceptions (which we identify with a $^\ddag$), which need to be
validated in the future. I then re-ran the misinterpreters against the
chosen answers. I terminated when all the remaining wrong answers were
either (a) found in very few students (we found a gap between 23\% and
13\%, and hence took 14\% as the threshold), (b) difficult for us to
attribute to a misconception, or (c) appeared to me to be ``Lispy''
and hence not of broad interest.

@@

% \paragraph{Confirmed Misconceptions}

% The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.

% \paragraph{Added Misconceptions}

% The misinterpreters helped me find misinterpretations that I had
% overlooked. For instance, consider this program from \quizState:
% \begin{lstlisting}
% (defvar x 12)
% (deffun (f x)
%   (set! x 0))
% (f x)
% x
% \end{lstlisting}
% I had assumed the wrong answer
% \texttt{0} is caused by \miscon{CallByRef}.
% However, my misinterpreters made me realize it can also be explained
% by \miscon{FlatEnv}. This could also explain why I see a difference in
% error rate when the formal and actual parameter have the same
% name (as mentioned in \Cref{s:misconceptions-quiz:state}).

% Tutor's refined program sets and misinterpreters also helped me to
% identify programs that detect one of the misconceptions but not the
% other. Consider this new program
% % mut-vars.not_aliased_by_defvar_1 (29)
% \begin{lstlisting}
%   (defvar x 12)
%   (defvar y x)
%   (set! x 0)
%   x
%   y
% \end{lstlisting}
% 12\% of students asserted that the following program produces
% \texttt{0 0}, which suggest that they have the \textbf{BindByRef}
% misconception. Students with a \miscon{FlatEnv} misconception would
% have chosen \texttt{0 12}, which happens to be the correct answer.
% (1\% wrote \textit{depends on
% implementation}; 2\% asserted \texttt{error}; the remaining 85\%
% asserted \texttt{0 12}).

% The second program in \Cref{t:quizzes-data-fun} is already a good
% detector for \miscon{FlatEnv}. SMoL Tutor includes an alternative
% program that does not depend on equality checks.
% % scope.error_when_refer_to_undefined (9)
% \begin{lstlisting}
%   (deffun (addy x)
%     (defvar y 200)
%     (+ x y))
%   (+ (addy 2) y)
% \end{lstlisting}
% 4\% of students asserted that the following program produces
% \texttt{402}, which suggest that they have the \miscon{FlatEnv}
% misconception. Students with a \textbf{BindByRef} misconception would
% have chosen \texttt{error}, which happens to be the correct answer.
% (the remaining 96\% asserted \texttt{error}).

@@

Once I turned this into a misinterpreter, I found that it
unexpectedly captured the program-answer pair for the following
program from \quizFun---which should be an error, due to the
double-binding of \lstinline|x| in the same scope block---and the answer
\lstinline|2 0|:
\begin{lstlisting}
  (defvar x 0)
  (defvar y x)
  (defvar x 2)
  x
  y
\end{lstlisting}
Previously, I had interpreted this only as \miscon{DefOrSet}, because
students had stated that the second \lstinline|(defvar x ....)|
``mutates'' or ``redefines'' \lstinline|x|. This is reminiscent of the
behavior of languages like Python, which use the same syntax both for
binding new variables and for mutating existing ones.

@@

\paragraph{Confirmed Misconceptions}

The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.

% \paragraph{Summary}

% To summarize, I used the SMoL Tutor, with an expanded set of
% programs, to investigate student misconceptions. I implemented the
% idea of misinterpreters to help me properly classify student
% performance. I were able to reconfirm all the previously identified
% misconceptions, refine some of them, and also identify potential new
% ones (that need further investigation).

\section{What Pedagogic Techniques are Effective?}%
\label{s:pedagogic-techniques}

@@

\section{The SMoL Tutor}%
\label{s:tutor}

% So far, I have focused on identifying problems. As noted earlier
% (\Cref{s:smol-quizzes}), I provided students with refutation texts
% after the quizzes, but it is unclear to what extent students read,
% understood, or internalized these. I also wanted to determine whether
% other populations run into these issues, and it was unclear whether we
% would be able to administer passive quizzes to them.

% Furthermore, while the quizzes may be a useful diagnostic, my goal is
% not only to find faults but to improve the understanding of basic
% programming language behavior. I believe (and presumably so does
% anyone else who writes a formal semantics!)\ that an understanding of
% these basic program behaviors is important, and even more so when it
% comes to understanding concurrency, ownership, and other advanced
% features. Even more simply, misunderstanding these clearly impacts
% development and debugging time.

% In response, I decided to create a \emph{tutor}: the SMoL Tutor. It is built
% around my quiz instruments and the detected misconceptions. We
% describe the Tutor from a user's perspective in \Cref{s:tutor}, and
% explain how it was populated from the
% SMoL Quizzes in \Cref{s:tutor-refine-program-set}.
% I then discuss what
% I learned from its data (\Cref{s:misconceptions-tutor}), and finally
% evaluate its effectiveness as a \emph{tutor} (\Cref{s:effect-tutor}).

% \subsection{The User Experience}

@@

% Each tutorial consists of a sequence of questions. Most questions in
% the Tutor are \newterm{interpreting questions}.\footnote{The Tutor
% also asked students to perform some other activities, like showing
% programs and asking for their heap content, which I do not cover in
% this paper.} These questions (illustrated in
% \Cref{f:smol-tutor-example}) are versions of the SMoL Quizzes
% (obtained by the process described in
% \Cref{s:tutor-refine-program-set}). In each of these questions,
% students are shown a program and asked to predict the program's
% running result(s) by answering an MCQ (with an ``Other'' option).
% After making a choice, students receive feedback. If a student chooses
% incorrectly, they are (a) given an explanation that is \emph{based on
% the misconception associated with that wrong answer} (or a generic
% one, if there is not a specific misconception), and then (b) asked to
% answer a second question:
% \begin{enumerate}

%   \item The second question is semantically the same as the first, but
%         with superficial changes (e.g., variable names, constants, and operators are
%         changed) so that the student cannot immediately
%         guess the answer.

%   \item Instead of multiple-choice, students must \emph{type} the answer
%         into a text box. (The Tutor normalizes text to accommodate variations.)
%         This is
%         intentional. First, I wanted to force reflection on the
%         explanation just given, whereas with an
%         MCQ, students could just guess. Second, I felt that students would
%         find typing more onerous than clicking. In case students
%         had just guessed on a question, my hope is that the penalty of having to
%         type means, on \emph{subsequent} tasks, they
%         would be more likely to pause and reflect before choosing.

@@

% Of the modules, \textbf{local} is the least portable across
% languages. While local binding is of course present in other
% languages, this is mostly covered using nested \lstinline|defvar|s in
% earlier modules, starting with \textbf{scope}. The distinction central
% to this module (between three local-binding constructs with slightly
% different scopes) is primarily a focus of Lispy languages. Therefore,
% I exclude this module from my analysis and instruments, and indeed
% plan to deprecate the module in future versions.

% \Cref{s:smol-quizzes} discusses the partial randomization employed by
% the SMoL Quizzes. In contrast, the SMoL Tutor does \emph{not}
% randomize question order. This is because the Tutor consists of more
% than just questions: it also has explanatory text. This text is based
% on the preceding problems. Authoring it is therefore somewhat like
% writing a textbook, with complex dependencies that can easily be
% broken.

\section{How Effective Is the Tutor?}
\label{s:effect-tutor}

@@

\section{How Generalizable is My Observations?}
\label{s:generality-tutor}

@@

The processes described in
\Cref{s:find-misconceptions-with-quizius,s:find-misconceptions-with-tutor}
have been conducted with SMoL's Lispy syntax (\Cref{s:smol}). This can
potentially introduce biases to the found misconceptions. For example,
when using SMoL Quizzes, I found that some students think the following program
produces \lstinline|'(4 5)| rather than \lstinline|'(1 2)|.
\begin{lstlisting}
  (filter (lambda (n) (> 3 n)) '(1 2 3 4 5))
\end{lstlisting}
This checks whether $\verb|3| > \verb|n|$, but those students
presumably vocalized it as ``greater than \lstinline|3|,
\lstinline|n|?''. Although I have tried to remove programs like this
from the tutor, there is still a risk of having biases in student
errors. This semester (Fall 2023) I am already deploying a newer SMoL
Tutor at Brown. This new tutor is able to display programs in
JavaScript and Python, in addition to the Lispy syntax. I plan to
compare in Spring 2024 the new tutor data with the current results,
i.e., a comparison between uni-syntax tutor and multi-syntax tutor.

>>>>

@@

To overcome such bias, I analyze a University~1 dataset produced by a
two-year process to find (a) student-written programs that tend to
trip up students and (b) incorrect responses to those programs using a tool for the purpose
(\Cref{s:quizius}). I have been cleaning up those programs
(\Cref{s:collating-problems}) and confirming that
these programs are still tricky in the same way.
\Cref{s:misconceptions-tutor} presents the final list of
misconceptions found with this process.

@@

\section{Preliminary Results}

@@

As said in \Cref{s:plan-rq1}, my main solution to the expert blind spot is to use the Quizius system
\citep{saarinenHarnessingWisdomClasses2019}. In contrast to the very
heavyweight process (involving a lot of expert time) that is generally
used to create a concept inventory, Quizius uses a lightweight,
interactive approach to obtain fairly comparable data, which an expert
can then shape into a quality instrument.

@@

\subsection{Collating Problems}
\label{s:collating-problems}

@@

  \item I felt the existing programs were insufficient to cover the
        ideas I wanted students to work through. For instance, I found no program that aliases vectors using \lstinline|defvar|, so we
        added this program:

@@

\end{itemize}

\subsection{Preliminary List of Misconceptions}%
\label{s:misconceptions-tutor}

@@

I iteratively created our final catalog of misconceptions (from the
perspective of the data in this paper). I started with misinterpreters
representing the misconceptions for which we
have reasonable validation (due to the prose in Quizius), so I call
them \emph{grounded} misconceptions. I then looked at wrong answers
not covered by these but chosen by students, and did my best to
distill these into misconceptions. These are \emph{surmised}
misconceptions (which we identify with a $^\ddag$), which need to be
validated in the future. I then re-ran the misinterpreters against the
chosen answers. I terminated when all the remaining wrong answers were
either (a) found in very few students (we found a gap between 23\% and
13\%, and hence took 14\% as the threshold), (b) difficult for us to
attribute to a misconception, or (c) appeared to me to be ``Lispy''
and hence not of broad interest.

@@

@@

Once I turned this into a misinterpreter, I found that it
unexpectedly captured the program-answer pair for the following
program --- which should be an error, due to the
double-binding of \lstinline|x| in the same scope block---and the answer
\lstinline|2 0|:
\begin{lstlisting}
  (defvar x 0)
  (defvar y x)
  (defvar x 2)
  x
  y
\end{lstlisting}
Previously, I had interpreted this only as \miscon{DefOrSet}, because
students had stated that the second \lstinline|(defvar x ....)|
``mutates'' or ``redefines'' \lstinline|x|. This is reminiscent of the
behavior of languages like Python, which use the same syntax both for
binding new variables and for mutating existing ones.

@@

\subsection{What Pedagogic Techniques are Effective?}%
\label{s:pedagogic-techniques}

@@

\subsection{The SMoL Tutor}%
\label{s:tutor}

@@

@@

\subsection{How Effective Is the Tutor?}
\label{s:effect-tutor}

@@

\subsection{How Generalizable are My Results?}
\label{s:generality-tutor}

@@

The processes described in
\Cref{s:find-misconceptions-with-quizius,s:find-misconceptions-with-tutor}
have been conducted with SMoL's Lispy syntax (\Cref{s:smol}). This can
potentially introduce biases to the found misconceptions. For example,
I found that some students think the following program
produces \lstinline|'(4 5)| rather than \lstinline|'(1 2)|.
\begin{lstlisting}
  (filter (lambda (n) (> 3 n)) '(1 2 3 4 5))
\end{lstlisting}
This checks whether $\verb|3| > \verb|n|$, but those students
presumably vocalized it as ``greater than \lstinline|3|,
\lstinline|n|?''. Although I have tried to remove programs like this
from the tutor, there is still a risk of having biases in student
errors. This semester (Fall 2023) I am already deploying a newer SMoL
Tutor at Brown. This new tutor is able to display programs in
JavaScript and Python, in addition to the Lispy syntax. I plan to
compare in Spring 2024 the new tutor data with the current results,
i.e., a comparison between uni-syntax tutor and multi-syntax tutor.
